Cyclone: High Availability for Persistent Key Value Stores

### 2018-02-07 paper/paper.tex
More tweaks.
Minor tweaks.

\author{{\rm Amitabha Roy} \\ amitabha.roy@gmail.com
	     \and {\rm Subramanya R. Dulloor} \\ subramanya.r.dulloor@intel.com}

@@

\section{Storage}
\label{sec:storage}
The storage bottleneck for the write ahead log in key value stores
such as Rocksdb is - surprisingly - not the storage media (i.e. SSD)
itself but rather the interface (i.e block IO). This is a particular
pain point for small writes.  For example,
Facebook~\cite{fb_workload_analysis} reports that 90\% of the space
allocated in key value stores is for data items under 500 bytes in
size.  The SSD used in Figure~\ref{fig:problem}, we achieve 50
Kops/sec to the SSD being limited to a 20us latency on the PCIe
interface. This translates to 25 MB/s for 500 byte key-value
pairs. The SSD is capable of achieving over 1 GB/s of sequential write
bandwidth to an append only log. Clearly the SSDs performance is not
the problem, but rather the block IO interface that imposes the round
trip latency \emph{regardless of the size of the update}.

@@

We further reduce work per-packet for the \emph{loaded} case using
batching - treating a whole sequence of client commands as a single
RAFT log entry, while avoiding any copies to group these packets
together. We do this by making use of a DPDK feature that allows
chaining received packet mbufs together and treating the chain as a
single packet.  Finally, we pipline the RAFT protocol state machine by
contnuing to transmit log entries to follower replicas without waiting
for acknowledgments. The RAFT protocol allows tracking multiple log
entries in flight via their entry index in the log and this allows us
to keep the link busy without being limited by the round trip time.
At follower replicas we also receive packets containing log entries in
batches and send back a single acknowledgment for the whole batch.

@@

\subsection{Ganged Operations}
Using multiple instances of RAFT leads to a serious problem with
batched operations, since we need to split up the batch into a
mini-batch for each of the physical logs. There is no guarantee that
the request for a mini-batch will succeed breaking the atomicity
requirement for batched updates. The obvious and simple solution is to
do two phase commit across the logs, ensuring all or nothing
semantics. However this was unacceptable to some use-cases of Cyclone
as it requires ensuring that an external co-ordinator implementing two
phase commit does not itself fail. This requires running a separate
replicated service - with its attendant resource and reliability
headaches - too high a price to pay for the performance benefits of
Cyclone.  We therefore had to come up with a solution that did not
require two phase commit.

@@

\section{Conclusion}
Cyclone shows how one can leverage a small amount of non-volatile
memory to address two fundamental difficulties in adding high
availability to persistent key value stores. First, Cyclone avoids
paying the cost of the block IO interface for every update
operation. In turn, this enables us to apply software packet
switching techniques and concurrency to improve the performance of
replication, and making full use of available resources when
replicating log entries over the network. It enables deployments of
Rocksdb with native performance without any compromise on availability.
Cyclone is available as open source software at:\\
\url{https://github.com/sdulloor/cyclone/}
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

\author{{\rm Amitabha Roy} \\ amitabha.roy@gmail.com \thanks{Work done while the author was at Intel Labs}
	     \and {\rm Subramanya R. Dulloor} \\ subramanya.r.dulloor@intel.com}

@@

\section{Storage}
\label{sec:storage}
The storage bottleneck for the write ahead log in key value stores
such as Rocksdb is - surprisingly - not the storage media (i.e. SSD)
itself but rather the interface (i.e block IO). This is a particular
pain point for small writes.  For example,
Facebook~\cite{fb_workload_analysis} reports that 90\% of the space
allocated in key value stores is for data items under 500 bytes in
size. For the SSD used in Figure~\ref{fig:problem}, we achieve 50
Kops/sec to the SSD due to being limited by the 20us latency on the PCIe
interface. This translates to 25 MB/s for 500 byte key-value
pairs. The SSD is capable of achieving over 1 GB/s of sequential write
bandwidth to an append only log. Clearly the SSDs performance is not
the problem, but rather the block IO interface that imposes the round
trip latency \emph{regardless of the size of the update}.

@@

We further reduce work per-packet for the \emph{loaded} case using
batching - treating a whole sequence of client commands as a single
RAFT log entry, while avoiding any copies to group these packets
together. We do this by making use of a DPDK feature that allows
chaining received packet mbufs together and treating the chain as a
single packet.  Finally, we pipeline the RAFT protocol state machine by
continuing to transmit log entries to follower replicas without waiting
for acknowledgments. The RAFT protocol allows tracking multiple log
entries in flight via their entry index in the log and this allows us
to keep the link busy without being limited by the round trip time.
At follower replicas we also receive packets containing log entries in
batches and send back a single acknowledgment for the whole batch.

@@

\subsection{Ganged Operations}
Using multiple instances of RAFT leads to a serious problem with
batched operations, since we need to split up the batch into a
mini-batch for each of the physical logs. There is no guarantee that
the request for a mini-batch will succeed breaking the atomicity
requirement for batched updates. The obvious and simple solution is to
do two phase commit across the logs, ensuring all or nothing
semantics. However this was unacceptable for some use-cases of Cyclone
as it requires ensuring that an external co-ordinator implementing two
phase commit does not itself fail. This requires running a separate
replicated service - with its attendant resource and reliability
headaches - too high a price to pay for the performance benefits of
Cyclone.  We therefore had to come up with a solution that did not
require two phase commit.

@@

\section{Conclusion}
Cyclone shows how one can leverage a small amount of non-volatile
memory to address fundamental difficulties in adding high
availability to persistent key value stores. Cyclone avoids
paying the cost of the block IO interface for every update
operation. In turn, this enables us to apply software packet
switching techniques and concurrency to improve the performance of
replication, and make full use of available resources when
replicating log entries over the network. It enables deployments of
Rocksdb with native performance without any compromise on availability.
Cyclone is available as open source software at:\\
\url{https://github.com/sdulloor/cyclone/}
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
